---
title: "Data Management Final Project"
author: "Tom Flachaire & Raphael Paillard"
format: pdf
---

## Project description

### Github link

Please find the github link [here](https://github.com/RaphP22/final_project).

### Sources links

With this link, we downloaded : [here](https://www.insee.fr/fr/statistiques/7624538?sommaire=7624746&utm_)

- P3D – Indicateurs généraux de la population par département et région - Séries depuis 1990
- P1D – Évolution générale de la situation démographique par département et région - Séries depuis 1990
- P2D – Mortalité infantile par département et région - Séries depuis 1990.

On the Data DREES website : 
[here](https://data.drees.solidarites-sante.gouv.fr/explore/dataset/la-demographie-des-professionnels-de-sante-depuis-2012/information/)

We downloaded data on midwife demographics from 2012 to 2025.

And with this link : https://www.insee.fr/fr/statistiques/6036902?sommaire=6036904
We downloaded the data about median income and poverty rates per departments.

Then, we moved them to csv files only keeping the year 2019, the last year available before COVID, which might affect the result.

### Description of the sources and important variables

The data comes from French sources, the INSEE (National Institute of Statistics and Economic Studies) and the DREES (Directorate for Research, Studies, Evaluation, and Statistics).

Infant Mortality provides department-level statistics on the number of deaths of children under one year old. These data, produced by INSEE and Santé publique France, help analyze regional disparities in infant health.

Population Indicators include information on fertility, average maternal age at childbirth, and the proportion of births outside marriage, broken down by department and year.

Demographic Situation contains data on marriages, population counts, nuptiality rates, birth rates, and natural population growth by department and year. 

Midwives RPPS 2012–2025 tracks the number and density of registered midwives across French departments.

### Basic information about each data file

```{r echo=FALSE}
#| message: false
#| warning: false
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
library(here)
library(stringr)
library(tidyr)
library(lmtest)
```

```{r echo=FALSE}
get_data_info <- function(file_path) {
  data <- read.csv2(file_path)
  
  cols_clean <- colnames(data)[!grepl("^X", colnames(data))]
  
  data.frame(
    Fichier = basename(file_path),
    Lignes = nrow(data),
    Colonnes = ncol(data),
    Colonnes_principales = paste(cols_clean, collapse = ", ")
  )
}

data_files <- list.files("data", pattern = "\\.csv$", full.names = TRUE)
data_info_df <- do.call(rbind, lapply(data_files, get_data_info))

data_info_df <- data_info_df %>%
  mutate(
    Colonnes_principales = case_when(
      Fichier == "indicateurs_population.csv" ~ "Indicateurs généraux de la population",
      Fichier == "mortalite_infantile.csv" ~ "Mortalité infantile",
      Fichier == "Sages-femmes_densité.csv" ~ "Densité des sages-femmes par département (2012–2025)",
      Fichier == "Sages-femmes_effectif.csv" ~ "Effectifs des sages-femmes par département (2012–2025)",
      Fichier == "situation_demographique.csv" ~ "Situation démographique générale",
      TRUE ~ Colonnes_principales
    )
  )

kable(
  data_info_df,
  caption = "Résumé des fichiers de données"
)
```

### Research Question

Why does infant mortality vary so much across French departments?

This project examines the factors that can explain variations in infant mortality rates across French departments. We will look at the role of demographic characteristics and access to maternal healthcare. While France performs relatively well in terms of infant health at the national level, there are huge territorial disparities. Some departments consistently have higher infant mortality rates than others, raising questions about the causes of these differences.

Demographic indicators such as fertility rates, maternal age at childbirth, and the proportion of births outside marriage—offer explains the social and economic contexts in which children are born. These factors may influence early childhood health risks.

A key aspect of this research is the distribution of midwives across departments. Midwives play a important role in premature deaths. 

The project aims to identify the factors most closely associated with variations in infant mortality. The objective is not only to describe regional disparities but also to understand their causes. 

We only kept data from 2019, the most recent year available before covid that could influence our results. 

### Data Preparation and Cleaning

```{r echo=FALSE}
#| message: false
#| warning: false
infant_mortality <- read_csv2(
  here("data/mortalite_infantile.csv"),
  skip = 4
) %>%
 mutate(
# Convert all numeric columns to numeric type since R doesn't necessarly understand it at first.
    `Nés vivants (a)` = parse_number(`Nés vivants (a)`, locale = locale(decimal_mark = ",", grouping_mark = " ")),
    `Décédés de moins d'un an (b)` = parse_number(`Décédés de moins d'un an (b)`, locale = locale(decimal_mark = ",", grouping_mark = " ")),
    `Décédés de moins de 7 jours (b)` = parse_number(`Décédés de moins de 7 jours (b)`, locale = locale(decimal_mark = ",", grouping_mark = " ")),
    `Décédés de moins de 28 jours (b)` = parse_number(`Décédés de moins de 28 jours (b)`, locale = locale(decimal_mark = ",", grouping_mark = " ")),
    `Infantile (c)` = parse_number(`Infantile (c)`, locale = locale(decimal_mark = ",", grouping_mark = " ")),
    `Néonatale (d)` = parse_number(`Néonatale (d)`, locale = locale(decimal_mark = ",", grouping_mark = " "))
  ) %>%
  filter(!is.na(`Décédés de moins d'un an (b)`))  

# We are deleting the lines with no values.

# we exclude regional total since we care about departments.
infant_mortality <- infant_mortality %>%
  filter(
    !grepl("France|métropole|Sous-total|Outre Mer", `Départements et régions`, ignore.case = TRUE),
    !`Départements et régions` %in% c("Ile-de-France", "Centre-Val de Loire", 
                                       "Bourgogne-Franche-Comté", "Normandie",
                                       "Hauts-de-France", "Grand Est", 
                                       "Pays de la Loire", "Bretagne",
                                       "Nouvelle-Aquitaine", "Occitanie",
                                       "Auvergne-Rhône-Alpes", "Provence-Alpes-Côte d'Azur",
                                       "Corse")
  )

```

We imported the infant mortality dataset from a CSV file using the `read_csv2()` function. 
The file contained data in the first four rows (title, year, and empty lines), 
which we skipped using the `skip = 4` parameter to start reading from the actual column headers. We filtered out rows containing NA values in the variables. We excluded regional totals and national aggregates to ensured we only kept individual department-level observations. 

All columns were initially imported as character type due to formatting issues so we converted the mortality count variable to numeric using `parse_number`.

Now let's clean the other databases.

For indicateurs_generaux database :

```{r echo=FALSE}
#| message: false
#| warning: false
indicateurs_generaux <- read_csv2(
  here("data/indicateurs_population.csv"),
  skip = 3
) %>%
  filter(!is.na(`Indicateur conjoncturel de fécondité (a)`))

# Keep only departments (exclude regional totals)
indicateurs_generaux <- indicateurs_generaux %>%
  filter(
    !grepl("France|métropole|Sous-total|Outre Mer", `Départements et régions`, ignore.case = TRUE),
    !`Départements et régions` %in% c("Ile-de-France", "Centre-Val de Loire", 
                                       "Bourgogne-Franche-Comté", "Normandie",
                                       "Hauts-de-France", "Grand Est", 
                                       "Pays de la Loire", "Bretagne",
                                       "Nouvelle-Aquitaine", "Occitanie",
                                       "Auvergne-Rhône-Alpes", "Provence-Alpes-Côte d'Azur",
                                       "Corse", "Départements et régions d'Outre Mer")
  )
```

For situation_demographique database :

```{r echo=FALSE}
#| message: false
#| warning: false
situation_demo <- read_csv2(
  here("data/situation_demographique.csv"),
  skip = 4
) %>%
  filter(!is.na(`Population moyenne`)) 

# Keep only departments (exclude regional totals)
situation_demo <- situation_demo %>%
  filter(
    !grepl("France|métropole|Sous-total|Outre Mer|Autres", `Départements et régions`, ignore.case = TRUE),
    !`Départements et régions` %in% c("Ile-de-France", "Centre-Val de Loire", 
                                       "Bourgogne-Franche-Comté", "Normandie",
                                       "Hauts-de-France", "Grand Est", 
                                       "Pays de la Loire", "Bretagne",
                                       "Nouvelle-Aquitaine", "Occitanie",
                                       "Auvergne-Rhône-Alpes", "Provence-Alpes-Côte d'Azur",
                                       "Corse", "Départements et régions d'Outre Mer")
  )
situation_demo <- situation_demo %>%
  mutate(
    across(
      .cols = !all_of("Départements et régions") & where(is.character),
      ~ parse_number(.x, locale = locale(decimal_mark = ",", grouping_mark = " "))
    )
  )
```

Now we do the data cleaning and joining for the DRESS data about midwives :

```{r echo=FALSE}

midwives_workforce_complete <- read.csv2(here("data/Sages-femmes_effectif.csv"), stringsAsFactors = FALSE)

# We only keep year 2019. And we don't care about the ages, the sector of activities and the gender of our midwives.

midwives_workforce <- midwives_workforce_complete %>%
  filter(
    secteur_activite == "00-Ensemble",
    exercice == "0-Ensemble",
    tranche_age == "00-Ensemble",
    sexe == "0-Ensemble"
  ) %>%
  select(departement, effectif_2019)


# We delete the "Ensemble ligns that we won't use later.
midwives_workforce <- midwives_workforce %>%
  filter(!grepl("Ensemble", departement))

# We clean the name of the departments which was "075-PARIS" because in our other datasets it's only "Paris".
  
midwives_workforce <- midwives_workforce %>%
  mutate(departement = sub("^\\d+\\s*-", "", departement))
```


```{r echo=FALSE}

# we do exactly the same for the midwives density.
midwives_density_complete <- read.csv2(here("data/Sages-femmes_densité.csv"), stringsAsFactors = FALSE)

midwives_density <- midwives_density_complete %>%
  filter(
    secteur_activite == "00-Ensemble",
    exercice == "0-Ensemble",
    tranche_age == "00-Ensemble",
    sexe == "0-Ensemble"
  ) %>%
  select(departement, densite_2019)

midwives_density <- midwives_density %>%
  filter(!grepl("Ensemble", departement))

midwives_density <- midwives_density %>%
  mutate(departement = sub("^\\d+\\s*-", "", departement))
```


```{r echo=FALSE}
# We join the 2 datasets to have only one for midwives : 

midwives <- inner_join(midwives_density, midwives_workforce, by = "departement")

midwives <- midwives %>%
  select(departement, effectif_2019, densite_2019)
```

We import a dataset about the income and poverty rate per department in 2019.

```{r echo=FALSE}

incomes <- read.csv2(
  file = here("data/revenus_2019.csv"),
  skip = 4,
  header = TRUE,
  stringsAsFactors = FALSE
  )

incomes <- incomes[-c(1, 100, 101, 102, 103), 1:9 ]

cols_num <- 3:ncol(incomes)

incomes[cols_num] <- lapply(incomes[cols_num], function(x) {
  as.numeric(gsub(",", ".", x))
})

## We don't keep the first 4th lign and neither the 6th because it's the 5th lign which has the name of variables we can understand. We dont keep the colomns after the 8th, it's indicators that we won't use.

incomes <- incomes %>%
  rename(
    'Taux de pauvreté' = `Taux.de.pauvreté.Ensemble....`,
    'Libellé géographique' = `Libellé.géographique`,
    num_households = `Nombre.de.ménages.fiscaux`,
    num_people = `Nombre.de.personnes.dans.les.ménages.fiscaux`,
    median_income = `Médiane.du.niveau.vie....`,
    pct_households_taxed = `Part.des.ménages.fiscaux.imposés....`,
    pct_poverty_under30 = `Taux.de.pauvreté.moins.de.30.ans....`,
    pct_poverty_30_39 = `Taux.de.pauvreté.30.à.39.ans.....`
  )

```


Sorting the datasets in alphabetical order and adding a number to each departement.


```{r echo=FALSE}

midwives <-midwives%>%
  filter(!grepl(" COM ou Etranger",departement,ignore.case = TRUE))

infant_mortality <-infant_mortality%>%
  filter(!grepl("Autres",`Départements et régions`,ignore.case = TRUE))





midwives$departement_clean <- midwives$departement

midwives$departement_clean[midwives$departement == "02A -Corse-du-Sud"] <- "Corse-du-Sud"
midwives$departement_clean[midwives$departement == "02B -Haute-Corse"] <- "Haute-Corse"

midwives$departement_clean[midwives$departement == " Martinique"] <- "Martinique"



midwives <- midwives[order(midwives$departement_clean), ]


midwives$number <- match(midwives$departement_clean, unique(midwives$departement_clean))





incomes <- incomes[order(incomes$'Libellé géographique'), ]

incomes$number <- 1:nrow(incomes)


indicateurs_generaux <- indicateurs_generaux[order(indicateurs_generaux$`Départements et régions`), ]

indicateurs_generaux$number <- 1:nrow(indicateurs_generaux)



infant_mortality <- infant_mortality[order(infant_mortality$`Départements et régions`), ]

infant_mortality$number <- 1:nrow(infant_mortality)



situation_demo <- situation_demo[order(situation_demo$`Départements et régions`), ]

```

### Main target variable

To illustrate our infant mortality variable, we will show the difference between top 5 and bottom 5 departments. Showing all 101 departments in one graph would be ugly.

```{r echo=FALSE}

# this is how we isolate the top and bottom 5

top5 <- infant_mortality %>%
  arrange(desc(`Décédés de moins d'un an (b)`)) %>%
  slice(1:5)

bottom5 <- infant_mortality %>%
  arrange(`Décédés de moins d'un an (b)`) %>%
  slice(1:5)

top_bottom <- bind_rows(
  top5 %>% mutate(categorie = "Plus élevé"),
  bottom5 %>% mutate(categorie = "Plus faible")
)

# and the graphical representation
ggplot(top_bottom, 
       aes(x = reorder(`Départements et régions`, `Décédés de moins d'un an (b)`), 
           y = `Décédés de moins d'un an (b)`,
           fill = categorie)) +
  geom_col() +
  geom_text(aes(label = `Décédés de moins d'un an (b)`), 
            hjust = -0.2, size = 3.5) +  
  scale_fill_manual(values = c("Plus élevé" = "coral", "Plus faible" = "steelblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +  # Plus d'espace à droite
  coord_flip() +
  labs(title = "Death of less than 1y per department (2019)",
       subtitle = "top 5 and bottom 5",
       y = "Number of deaths", 
       x = NULL,
       fill = NULL) +
  theme_minimal() +
  theme(legend.position = "top")
```

We can see on the graph that there is a huge disparity between departments in infant mortality. For example, the highest deaths are in Seine Saint Denis (148) whereas in Cantal there weren't any death.

## Descriptive Statistics

### Poverty and Infant Mortality

Socio-economic conditions are widely recognized as key determinants of early childhood health. In particular, poverty may limit access to healthcare, increase exposure to health risks, and negatively affect pregnancy outcomes. This section examines the relationship between poverty rates and infant mortality across French departments in 2019.

```{r echo=FALSE}
#| message: false
#| warning: false
#| fig.width: 9
#| fig.height: 6

ggplot(incomes, aes(
x = reorder(`Libellé géographique`, `Taux de pauvreté`),
y = `Taux de pauvreté`
)) +
geom_col(fill = "steelblue") +
coord_flip() +
labs(
title = "Poverty rate by French department (2019)",
x = "Department",
y = "Poverty rate (%)",
caption = "Source: INSEE"
) +
theme_minimal()

```
In overseas departments, poverty rates are significantly higher, while in mainland France, poverty is more homogeneous and generally lower, with a few exceptions like Seine-Saint-Denis. This reflects economic disparities in poverty rate, and that is important to notice before analysing the link between poverty and infant mortality.

```{r echo=FALSE}
#| message: false
#| warning: false
#| fig.width: 7
#| fig.height: 5

ggplot() +
geom_density(
data = incomes,
aes(x = `Taux de pauvreté`, fill = "Poverty rate"),
alpha = 0.5
) +
geom_density(
data = infant_mortality,
aes(x = `Infantile (c)`, fill = "Infant mortality"),
alpha = 0.5
) +
labs(
title = "Distribution of poverty and infant mortality across departments (2019)",
x = "Value (poverty % / infant mortality ‰)",
y = "Density",
fill = ""
) +
theme_minimal()

```
Description of the Graph
- The x-axis represents the values of poverty rates (in percentage) and infant mortality rates (in per thousand).
- The y-axis indicates the density, showing the frequency of these values across the departments.

The red curve shows the infant mortality rate, which stays mostly low in many departments, around 4%. 
The green curve shows the poverty rate, which is a bit more spread out. Some departments have very little poverty, while others have a lot, especially around 10-20%. This tells us that poverty is not the same everywhere—it changes a lot from one department to another.

```{r echo=FALSE}
#| echo: false
#| message: false
#| warning: false
#| fig.width: 8
#| fig.height: 5

# Créer un dataframe "long" pour ggplot en utilisant les backticks pour les noms de colonnes
df_long <- data.frame(
  Département = rep(infant_mortality$`Départements et régions`, 2),
  Valeur = c(
    infant_mortality$`Infantile (c)`,
    as.numeric(gsub(",", ".", incomes$`Taux.de.pauvreté.Ensemble....`))
  ),
  Indicateur = rep(
    c("Mortalité infantile (‰)", "Taux de pauvreté (%)"), 
    each = nrow(infant_mortality)
  )
)

ggplot(df_long, aes(x = Indicateur, y = Valeur, fill = Indicateur)) +
  geom_boxplot(alpha = 0.7, outlier.colour = "red", outlier.size = 2) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 1.5) +
  scale_fill_manual(values = c("steelblue", "coral")) +
  labs(
    title = "Distribution du taux de pauvreté et de la mortalité infantile (2019)",
    x = "Indicateur",
    y = "Valeur",
    caption = "Source : INSEE & DREES"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```
This boxplot provides a comparison of the distribution of infant mortality rates (‰) and poverty rates (%) across French departments in 2019. 

The blue boxplot shows that most departments have infant mortality rates clustered between approximately 2.5‰ and 4‰. The median rate is around 3.5‰. However, there are several outliers, with some departments experiencing rates as high as 10‰. These outliers suggest that certain areas face significantly higher infant mortality, potentially due to socio-economic factors or healthcare access issues.

The orange boxplot for poverty rates shows a similar result, ranging from nearly 2.5% to over 4%. The median poverty rate is around 3%. The presence of numerous outliers, especially on the higher end, indicates that some departments have substantially higher poverty rates. 

```{r echo=FALSE}
#| echo: false
#| message: false
#| warning: false

# Créer un vecteur propre pour le taux de pauvreté
poverty_rate <- as.numeric(gsub(",", ".", incomes$`Taux de pauvreté`))  # adapter selon ton colnames exact
infant_rate <- infant_mortality$`Infantile (c)`[1:length(poverty_rate)]  # garder même nombre de lignes

# Créer le data frame
plot_data <- data.frame(
  poverty_rate = poverty_rate,
  infant_mortality_rate = infant_rate
)

# Graphique
ggplot(plot_data, aes(x = poverty_rate, y = infant_mortality_rate)) +
  geom_point(color = "coral", size = 3, alpha = 0.6) +
  geom_smooth(method = "lm", color = "steelblue", fill = "lightblue", alpha = 0.2) +
  labs(
    title = "Poverty Rate vs Infant Mortality Rate (2019)",
    x = "Poverty Rate (%)",
    y = "Infant Mortality Rate (‰)"
  ) +
  theme_minimal()

```
Now we can link those two variables. This scatter plot clearly shows a positive correlation between poverty rate and infant mortality rate in French departments for 2019.
Most departments cluster in the lower-left corner, indicating lower poverty rates (below 20%) and lower infant mortality rates (below 5‰). However, as the poverty rate increases—especially beyond 40%—there is a noticeable upward trend in infant mortality rates, reaching up to 10‰ or more.
The outliers in the top-right corner (departments with poverty rates above 60% and infant mortality rates above 7.5‰) underlines areas where socio-economic challenges likely contribute to poorer health outcomes. This reinforces the idea that reducing poverty could significantly improve infant survival rates.


```{r echo=FALSE}
data_model <- data.frame(
  mortalite = infant_mortality$`Infantile (c)`,
  revenu = incomes$median_income,
  age_birth = indicateurs_generaux$`Âge moyen à l'accouchement (a)`,
  pauvrete = incomes$`Taux de pauvreté`,
  hors_mariage = indicateurs_generaux$`Proportion de naissances hors mariage (a)`
) %>% na.omit()

# Correlation table

kable(round(cor(data_model), 3))
```

The analysis reveals that poverty has a strong positive correlation (0.584) with infant mortality, confirming that departments with higher poverty rates consistently exhibit worse infant health outcomes. Conversely, median income shows a moderate negative correlation (-0.461), suggesting that wealthier departments benefit from better healthcare access and resources. Maternal age and births outside marriage display only weak correlations (-0.072 and 0.208, respectively), indicating their limited impact compared to socio-economic factors like poverty and income.

```{r echo=FALSE}
# regression

model1 <- lm(mortalite ~ revenu + age_birth + pauvrete + hors_mariage, data = data_model)
summary(model1)
```
We ran a regression to see what really impacts infant mortality rates. Turns out, poverty is the big one—it’s statistically significant (p = 0.001) and has a positive effect. Basically, the higher the poverty rate in a department, the higher the infant mortality. For every 1% increase in poverty, infant mortality goes up by about 0.14‰. That’s a pretty clear sign that economic conditions matter a lot.
The other factors—like income, maternal age, and births outside marriage —don’t show a significant impact here. That was a bit surprising, especially since income seemed important in the correlation table, but I guess poverty is the stronger driver.
The model explains about 33% of the variation in infant mortality (Adjusted R-squared = 0.33), which shows that socio-economic factors play a big role. There’s probably more to the story, like access to healthcare or local policies, but this is a solid starting point. The low p-value (8.719e-09) confirms the overall significance of the model.

```{r echo=FALSE}
# Plot regression
par(mfrow = c(2, 2))
plot(model1)
```

Residuals vs Fitted:
The residuals are randomly scattered around zero, indicating that the model fits the data well without any obvious patterns. This suggests that the chosen functional form of the model is appropriate.

Q-Q Plot:
The residuals largely follow the theoretical quantile line, confirming that they are approximately normally distributed. Minor deviations at the tails may indicate some outliers, but they are not severe.

Scale-Location Plot:
The residuals display constant variance across fitted values, with no funnel shape visible. This supports the assumption of homoscedasticity, indicating no problematic heteroscedasticity in the model.

Residuals vs Leverage:
A few observations show high influence on the model, as indicated by their Cook’s distance. 

## Midwives and Infant Mortality

```{r echo=FALSE}
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

# Prepare infant mortality data
infant_data <- infant_mortality %>%
  mutate(
    departement = `Départements et régions`,
    infant_mortality_rate = `Infantile (c)`
  ) %>%
  select(departement, infant_mortality_rate)

# Prepare midwives data (already joined)
midwives_plot <- midwives %>%
  mutate(
    departement = departement,
    effectif = effectif_2019,
    densite = densite_2019
  ) %>%
  select(departement, effectif, densite)

# Join by department (only for plotting)
midwives_plot_data <- midwives_plot %>%
  inner_join(infant_data, by = "departement")

# Scatter plot: Midwife density vs Infant mortality
ggplot(midwives_plot_data, 
       aes(x = densite, y = infant_mortality_rate, size = effectif)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "coral", linetype = "dashed") +
  scale_size_continuous(name = "Number of midwives") +
  labs(
    title = "Midwives Density and Infant Mortality Across French Departments (2019)",
    x = "Midwives density (per 10,000 inhabitants)",
    y = "Infant mortality rate (‰)",
    caption = "Source: INSEE & DREES | Point size = total number of midwives"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )
```
This scatter plot illustrates the relationship between midwife density (represented by the size of the circles) and infant mortality rates in French departments.

The regression line shows a slight positive trend, suggesting that higher midwife density is unexpectedly associated with slightly higher infant mortality rates. This is counterintuitive.
The data points are widely scattered, indicating a weak relationship. Some departments with high midwife density still have high infant mortality rates, suggesting other influential factors (e.g., socio-economic conditions).


```{r echo=FALSE}
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 7

ggplot(midwives_plot_data, 
       aes(x = effectif, y = infant_mortality_rate, size = densite)) +
  geom_point(alpha = 0.7, color = "darkgreen") +
  geom_smooth(method = "lm", se = TRUE, color = "orange", linetype = "dashed") +
  scale_size_continuous(name = "Midwives effectifs") +
  labs(
    title = "Total Number of Midwives vs Infant Mortality Across French Departments (2019)",
    x = "Total number of midwives (effectif 2019)",
    y = "Infant mortality rate (‰)",
    caption = "Source: INSEE & DREES | Point size = midwives density per population"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )
```
Now let's illustrate the relationship between midwife effectif (represented by the size of the circles) and infant mortality rates in French departments.

The dashed regression line indicates a slight positive trend, suggesting that departments with more midwives do not necessarily have lower infant mortality rates. This implies that simply having more midwives may not directly translate to better infant health outcomes. There is no strong impact.

```{r echo=FALSE}

# Prepare dataset
midwives_plot_data <- midwives %>%
  select(departement, effectif_2019, densite_2019) %>%
  inner_join(
    infant_mortality %>%
      select(`Départements et régions`, `Nés vivants (a)`, `Décédés de moins d'un an (b)`) %>%
      mutate(infant_mortality_rate = (`Décédés de moins d'un an (b)` / `Nés vivants (a)`) * 1000),
    by = c("departement" = "Départements et régions")
  )

# Density plot of midwives
ggplot(midwives_plot_data, aes(x = densite_2019)) +
  geom_density(fill = "steelblue", alpha = 0.6) +
  labs(
    title = "Density of Midwives per Department (2019)",
    x = "Midwives Density",
    y = "Density"
  ) +
  theme_minimal()

```

Let's illustrates how midwives were distributed across different departments in France during 2019.
The majority of departments had a midwife density clustered between approximately 150 and 180, which appears to be the most common range. The curve peaks here, showing where most departments lie in terms of midwife availability.
This curve looks like a normal distribution, with a clear central peak and tails that gradually taper off toward the extreme values.

```{r echo=FALSE}
# Linear regression: Infant mortality ~ Midwives density
lm_density <- lm(infant_mortality_rate ~ densite_2019, data = midwives_plot_data)
summary(lm_density)
```

Finally, we examines the relationship between infant mortality rate and midwives density in 2019. Is there a real link ?
The coefficient for midwives density is 0.0144, and it is statistically significant (p-value = 0.00188), suggesting that midwives density has a positive, though modest, association with infant mortality rate in this model. However, the R-squared value of 0.1002 indicates that only about 10% of the variation in infant mortality rate is explained by midwives density, implying other factors likely play a more substantial role.

While the analysis reveals a counterintuitive positive association between midwives density and infant mortality rate, this does not imply a causal relationship. Many other important factors such as population size, poverty levels, access to healthcare, and every socioeconomic conditions likely influence infant mortality far more significantly.
Therefore, we cannot conclude that an increase in midwives density directly affects infant mortality without accounting for these broader contextual variables.


